Sediment transport capacity response to variations in water discharge in pressurized subglacial channels

### 2025-02-03 revised_v1/response_R1.tex

\textbf{General Comments}

\begin{itemize}

\item \textbf{In my view the paper is interesting and relevant, and hence worthy of publication. I have a couple
    of general observations, followed by a list of minor suggestions.
    My general observation relates to the structure of the main findings reported by the manuscript.
    In reading the Results and Discussion I was sometimes lost as to what I was learning that was new.
    I think this was mainly because the structure lacked a clear and logical progression from one
    general finding through progressively more specific findings. This may well stem an initial
    imperfect expression of the objectives.}

  \textit{We thank the Reviewer for their positive assessment of our work and acknowledge the structural and organizational issues they have highlighted.}

@@

\item  \textbf{Here, most of what is novel seems to fgall under Objective
    2 and not under Objective 1. Indeed, isn’t Objective 1 (‘to establish whether sub-seasonal water
    discharge can co-vary with sediment transport capacity in subglacial systems’) already established
    beyond reasonable doubt? To me, and I believe the introduction to this manuscript, it clearly
    ‘can’… (note the objective refers to ‘can it’, and not to ‘does it’ [to which the answer is also almost
    certainly ‘yes’ but perhaps with a little more room for maneuver]). I would recast the objectives
    to reflect better the new material presented in this manuscript (closer to objective 2 in fact, which
    I feel could readily be sub-divided). Having considered this, I feel the progression of the Results
    could be far more accessible to the reader and allow the manuscript to focus on a few really key
    points (which I think are currently slightly lost in the density of the presentation).
    Interpretation is presented within each Results section (e.g., see ‘due to… on L174). This
    continues throughout Results, and I would separate all of the explanation and hence
    interpretation out form the results.}

  \textit{We greatly appreciate this comment. After careful consideration, we have chosen to rephrase Objective One as:} ``to establish the hydrological conditions under which sediment transport capacity co-varies with water discharge in subglacial systems.''

\textit{In our view, this objective directs attention to the differences between the ALPINE and ICESHEET scenarios. It is also practical for evaluating the effects of water discharge smoothing, as discussed in the next point.}

\textit{We acknowledge the importance of removing interpretation from the results section and have made the necessary revisions. Additionally, we will moved Section 4.3, which contains the algebraic formulations, to the discussion section, as we believe that the pipe flow and steady-state R-channel scenarios serve as useful end members for interpreting the results.
}

@@

\item \textbf{There are some very important points in Results, which I would attach some key data to
    and make more explicit here and in the Conclusion and Abstract. I’m thinking for example of
    associating the (increased) hysteresis as a function of discharge with a headline \% change in
    sediment transport capacity. E.g., “thus, a more variable, but typical, Alpine Q record can
    transport up to *\% greater sediment per cumec than a less variable ice sheet-type Q” and/or “the
    greater variability in Q from Alpine than from ice sheet style Q can account by an offset of peak
    sediment discharge from peak meltwater discharge by up to * \% of the cycle involved”. At present,
    I feel the manuscript lacks this incisive illustration to bring the scale of the results home to the
    reader.}

  \textit{This is an excellent comment, and we have worked to provide a clearer explanation of hysteresis.}

\textit{One challenge with the approach recommended by the reviewer is that sediment transport capacity scales highly non-linearly with water discharge, particularly in subglacial systems, as demonstrated in Section 4.3. Consequently, findings across different hydrographs would not be directly comparable, even if they were scaled using a discharge quantity.}

\textit{Instead, we have added several paragraphs to Section 4.1 discussing the role of water discharge variability in hysteresis.}

@@

   \textit{ We will add the following sentence to the model implementation section:}
``We also note that the subglacial results apply to pressurize subglacial channels, not depressurized one that can occur below glacier termini \citep{perolo2018}''
\end{itemize}

\textbf{Specific Comments}

\begin{itemize}
\item \textbf{Title:}The title states the obvious from prior research and I would recast it to
  reflect more accurately the new and original contribution of this
  manuscript. Also, if the authors elect to not change the title I would change
  the existing wording to ‘transport capacities respond differently’.
  Grammatically, this is not clear given the two capacities noted earlier in the
  title, but stylistically I think ‘capacities respond’ reads better than does
  ‘capacity respond’. (I would change the whole thing anyway… Why not start
  by considering ‘Sediment transport capacity response to variations in water
  discharge in pressurized subglacial channels’)

  \textit{Excellent suggestion, title follows the reviewer's suggestion:} Sediment transport capacity response to variations in water discharge in pressurized subglacial channels

@@

\item \textbf{75-76 and 80-81} I’d remove summary of the results of this study from the Introduction. They
  are not yet established.

  \textit{These lines will be removed.}

\item \textbf{70-71} In phrasing objective 1 (in more than one place in the manuscript),
  shouldn’t the dependent variable come first? Thus, I would write it as ‘to
  establish whether sediment transport capacity can co-vary with water
  discharge in subglacial systems’.

  \textit{Given the discussion above, the text will read:} ``to establish the hydrological conditions where  sediment transport capacity co-varies with water discharge in subglacial systems''

@@

\item \textbf{79} Given the apparent relevance of the Alley et al. study, I feel it’s key findings
  should be presented in detail here to provide more direction to the specific
  objectives of this study.

  \textit{ We believe that this is discussed in Lines 51--54 in the introduction.}

\item \textbf{159} The parameter values are not purely random (but random within certain
  sensible ranges, no?).

  \textit{The text will now read:} ``with random parameter values in a range''

@@

\item \textbf{173} Change rather than evolution? (elsewhere too)

  \textit{We appreciate the comment. We believe that given the nature of subglacial channels as describe in Equation 3, evolution is a more accurate description. To us, ``change'' could include a modification for any number of reasons. However,
    ``evolution'' conotates the dependence on anticedent conditions which we believe to be more accurate in this case. }

\item \textbf{Fig 2 caption} Is it really an arbitrary y axis scale? If it is truly arbitrary, does it even need
  a scale range?

  \textit{We believe that description of the insets' axis is needed as these lines have the same axis in the main plots. However, their very different values mean that in makes sense to adjust values to examine their changes during the extreme events. The text will be modified slightly to read:} ``different y-axis ranges for subaerial and subglacial values.''

@@

\item \textbf{179-183} Is this not already established? If so, it should be presented clearly in the
  Introduction and developed as necessary here. It is also interpretation.

  \textit{ To the best of our knowledge this is the first time that this finding is being presented.
    We would welcome any citations or references that suggest otherwise. We believe that the content in this paragraph describes the response to water discharge in Figure 2. We agree that the reference to the ``Methods'' section, gave the impression of the interpretation. The paragraph will be adjusted slightly to read:}

@@

\item \textbf{205-6} Interpretation in Results

  \textit{Removed.}

\item \textbf{210-11} Interpretation in Results

  \textit{ Sentence will be changed to:}''Smaller values of channel factor $\beta$, creating low and broad channels where the channel width grows more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~4).''

@@

  \textit{Fixed.}

\item \textbf{Section 4.3} A lot of this reads more like methods and background modelling operation
  and/or equation terms than Results of the modelling. If it is all results then
  the key findings are difficult to pull out from the density of the presentation
  – perhaps some of it could be omitted or moved to supplementary?

  \textit{We have considered the comments here, and with the editors support, think it fits best here.
    In a previous version of the manuscript, it was included in the supplementary material. However, after those reviews, we believe that some misunderstandings emerged. We believe that Tables 2 and 3 in the section are valuable results in evaluating the sensitivity of sediment transport capacity to water discharge under different assumptions.
    The importance of these findings are discussed in lines xxx yyy.
  }

  \textit{We will modify the introductory text to read}: ``The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~4.3). 
Additionally, results demonstrate the impact of water discharge variability on sediment transport capacity.
Here, we compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient.''

@@

\item \textbf{319} I’m not sure ‘sporadic’ is the right word here. Please check you mean this.

  \textit{We will replace ``sporadic'' with ``variable''.}

\item \textbf{321-22} I see the point being made here but it’s not black or white. Surely there is
  some control exerted by discharge; it’s just that it is complicated by the
  additional forces in the subglacial scenario. The way this is presented here
  may be technically ok but the reader could be forgiven for coming away
  with the feeling that there is no solid relationship between Q and sediment
  transport capacity is these cases – which would not be accurate. I feel it
  would be particularly helpful if the effect could be quantified in some
  generalized or headline way in terms of the likely (or maximum if you
  prefer) effect on a typical alpine and/or ice sheet hydrograph. That figure
  could then also go into the Conclusions and Abstract.

  \textit{The sentence will read:} ``This hysteresis can limit water discharge as an indicator of sediment discharge capacity in these systems, especially when water discharge is highly variable and out of equilibrium with subglacial channel size.''

@@

\item \textbf{394} Surely the relationship can be characterized a little more usefully than
  simply to state that it is ‘incoherent’. This also relates to my point above
  relating to lines 321-22. Can this be characterized accurately or
  systematically so the reader has a little more to learn in terms of the nature
  and magnitude of the effects?

\textit{This paragraph will read:} ``The manuscript's first objective is to establish the conditions where water discharge covaries with subglacial sediment transport capacity.
In subglacial channels, the timing of peak water velocity and sediment transport capacity occurs before peak water discharge during a discharge event, due to evolving channel size.
In subaerial channels, the timing of peak sediment transport capacity and water discharge coincide.
Results here suggest that, even in a transport-limited subglacial system, an variable relationship between water and subglacial sediment discharge.
Water discharge variations in both Greenland Ice Sheet and Alpine are variable enough to cause this variable relationship.
This variable relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size, such as Antarctica with minimal surface melt input.
Results also suggest that water discharge records averaged over periods longer than $12$ hours from Alpine and Ice Sheet hydrographs show substantial impacts on subglacial sediment transport capacity characteristics.''

### 2025-02-03 revised_v1/response_editor.tex

\documentclass[11pt]{article}
\usepackage{lscape}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\setlength{\parindent}{0pt}
\usepackage [colon]{natbib}
\usepackage {comment}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=2.40cm]{geometry}
\usepackage{doi}
\usepackage{multirow}
\usepackage[T1, OT1]{fontenc}
\DeclareTextSymbolDefault{\dh}{T1}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{wrapfig}
% \linenumbers
\usepackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{float}
% \errorcontextlines=3
\usepackage[explicit]{titlesec}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[export]{adjustbox}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\title{}
\author{}

\begin{document}
Dear Editor,

We thank you for handling this manuscript and coordinating the reviews. We found both reviews very helpful in improving the manuscript.

@@

\vspace{.2cm}

Following Reviewer 1’s feedback, we have revised the first objective of the manuscript to better describe the conditions under which covariance between sediment transport capacity and water discharge may occur. To further develop this point, we conducted additional experiments using a series of smoothed hydrographs to better quantify the conditions that lead to covariance. Additionally, we have reorganized and revised the text in response to their comments.

At several points in the review, it was noted that a particular point had already been established. If this reflects a substantial oversight of the literature on our part, we sincerely appreciate the reviewer bringing it to our attention. However, we would be grateful if the reviewer could provide specific references in these instances to help us fully address their concerns.

@@

\vspace{.2cm}

Regarding Reviewer 2, we believe that their specific comments about the model have been addressed by generalizing the equations in question. We acknowledge their critique concerning the accumulation of water flow along the glacier’s length. However, we do not see a rationale suggesting that these findings would not hold for shorter channels or within a more complex (i.e., spatially distributed) model. To further emphasize these concerns, we have moved the ``Experiment Limitations'' discussion to a separate section following the ``Results'' section.

Once again, we sincerely appreciate your assistance with this manuscript and hope that you find our responses satisfactory.

### 2025-02-03 Delaneyetal2024.tex
changes to share

\begin{abstract}
  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with water velocity and channel width.
  In full subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
  However, in subglacial channels, water is pressurized by the ice above, and they grow in response to frictional heating of water flowing through them. 
  As a result, rapid changes in water discharge  mainly result in velocity variations, as the channel geometry evolves slowly.
  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice Sheet. 
  Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
  This hysteresis causes a highly variable relationship between sediment and water discharge in a transport-limited subglacial system.
  The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
  Reducing water discharge variability lessens the hysteresis effects, in some cases to the point where a covarying relationship between water discharge and sediment transport capacity may occur, similar to subaerial systems.
  A second set of numerical experiments shows that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
  Yet, results and  formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
  We present these results in the context of different formulations of the relationship between water discharge and sediment transport capacity based assumptions of subglacial and subaerial flow regimes.
  The implications of these findings help to evaluated sediment discharge from glaciers with different hydro-climatic forcings.
  These findings can improve the interpretation of sediment discharge records in glacierized catchments.
\end{abstract}

>>>>

\begin{abstract}
  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with water velocity and channel width.
  In full subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
  However, in subglacial channels, water is pressurized by the ice above, and they grow in response to frictional heating of water flowing through them. 
  As a result, rapid changes in water discharge  mainly result in velocity variations, as the channel geometry evolves slowly.
  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice Sheet. 
  Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
  This hysteresis causes a highly variable relationship between sediment and water discharge in a transport-limited subglacial system.
  The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
  Reducing water discharge variability by smoothing lessens the hysteresis effects, in some cases to the point where a covarying relationship between water discharge and sediment transport capacity may occur, similar to subaerial systems.
  A second set of numerical experiments shows that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
  Yet, results and  formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
  We present these results in the context of different formulations of the relationship between water discharge and sediment transport capacity based assumptions of subglacial and subaerial flow regimes.
  The implications of these findings help to evaluated sediment discharge from glaciers with different hydro-climatic forcings.
  These findings can improve the interpretation of sediment discharge records in glacierized catchments.
\end{abstract}
